---
title: "EDA"
author: "Gaurav Mishra"
date: "2026-02-22"
output: html_document
---
# Background 
- Each row = one QAL lead
- QAL = Qualified Account Lead 
- Someone showed interest and was qualified enough to enter the pipeline
- This dataset tracks marketing leads and what happened to them in the sales funnel


```{r}
# Install if needed
install.packages("tidyverse")
install.packages("janitor")
install.packages("lubridate")

# Load
library(tidyverse)
library(janitor)
library(lubridate)
```


```{r}
df <- read_csv("~/Downloads/QAL Performance for MSBA.csv")
```

## Structural Checks
- 16,812 rows and 14 columns 
```{r}
dim(df)
glimpse(df)
```
# Converting variable names to lowercase with underscores
```{r}
df <- df %>% clean_names()
names(df)
```
# Filtered out records with missing values

```{r}
df %>%
  count(next_stage_c, sort = TRUE)
```


```{r}
df_clean <- df %>%
  filter(!is.na(next_stage_c))

```

# Show how many rows were removed
```{r}
nrow(df) - nrow(df_clean)

```

# Baseline conversion rate
- 16-17%
```{r}
df_clean %>%
  summarise(
    total = n(),
    sqls = sum(next_stage_c == "SQL"),
    conversion_rate = sqls / total
  )

```
# Conversion by Marketing Channel

# Best performing channels --> highest quality leads 
- Direct/Inbound
- SEP
- Online Ads

```{r}
df_clean %>%
  group_by(last_tactic_campaign_channel) %>%
  summarise(
    total_leads = n(),
    sqls = sum(next_stage_c == "SQL"),
    conversion_rate = sqls / total_leads
  ) %>%
  arrange(desc(conversion_rate))
```

# Conversion by industry

# Highest SQL Conversion Rate
- Blood & biologics
- Medical devices
- Pharma & BioTech 
```{r}
df_clean %>%
  group_by(acct_target_industry) %>%
  summarise(
    total_leads = n(),
    sqls = sum(next_stage_c == "SQL"),
    conversion_rate = sqls / total_leads
  ) %>%
  arrange(desc(conversion_rate))

```

# Conversion by Account Tier 
- Small accounts convert best, medium, then large

```{r}
df_clean %>%
  group_by(acct_tier_rollup) %>%
  summarise(
    total_leads = n(),
    sqls = sum(next_stage_c == "SQL"),
    conversion_rate = sqls / total_leads
  ) %>%
  arrange(desc(conversion_rate))

```

# Industry x Tier
```{r}
df_clean %>%
  group_by(acct_target_industry, acct_tier_rollup) %>%
  summarise(
    total_leads = n(),
    sqls = sum(next_stage_c == "SQL"),
    conversion_rate = sqls / total_leads,
    .groups = "drop"
  ) %>%
  arrange(desc(conversion_rate))

```

# Time Trend
```{r}
library(lubridate)

df_clean %>%
  mutate(month = floor_date(qal_cohort_date, "month")) %>%
  group_by(month) %>%
  summarise(
    total_leads = n(),
    sqls = sum(next_stage_c == "SQL"),
    conversion_rate = sqls / total_leads
  ) %>%
  arrange(month)

```

# Visualization
```{r}
library(ggplot2)

df_clean %>%
  mutate(month = floor_date(qal_cohort_date, "month")) %>%
  group_by(month) %>%
  summarise(
    total_leads = n(),
    sqls = sum(next_stage_c == "SQL"),
    conversion_rate = sqls / total_leads
  ) %>%
  ggplot(aes(x = month, y = conversion_rate)) +
  geom_line(size = 1.2, color = "steelblue") +
  geom_point(size = 2) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    title = "Monthly SQL Conversion Rate",
    x = "Month",
    y = "Conversion Rate"
  ) +
  theme_minimal()

```
# Verify if spike is normal
```{r}
df_clean %>%
  mutate(month = floor_date(qal_cohort_date, "month")) %>%
  group_by(month) %>%
  summarise(total_leads = n()) %>%
  arrange(desc(month))
```


```{r}
df_clean %>%
  mutate(month = floor_date(qal_cohort_date, "month")) %>%
  group_by(month) %>%
  summarise(total_leads = n()) %>%
  ggplot(aes(month, total_leads)) +
  geom_line(linewidth = 1.2) +
  geom_point() +
  labs(
    title = "Monthly Lead Volume",
    x = "Month",
    y = "Total Leads"
  ) +
  theme_minimal()
```
# Overall SQL Rate
- Baseline conversion rate - 16 - 17%
```{r}
df_clean %>%
  mutate(is_sql = next_stage_c == "SQL") %>%
  count(is_sql) %>%
  mutate(percentage = n / sum(n))
```
# Territory conversion
- Japan has highest conversion rate but has a very small smaple size 
```{r}
df_clean %>%
  group_by(acct_territory_rollup) %>%
  summarise(
    total_leads = n(),
    sqls = sum(next_stage_c == "SQL"),
    conversion_rate = sqls / total_leads
  ) %>%
  arrange(desc(conversion_rate))
```

```{r}
library(scales)

heat_data <- df_clean %>%
  group_by(acct_target_industry, acct_tier_rollup) %>%
  summarise(
    conversion_rate = mean(next_stage_c == "SQL"),
    .groups = "drop"
  )

ggplot(heat_data, 
       aes(acct_tier_rollup, acct_target_industry, fill = conversion_rate)) +
  geom_tile() +
  scale_fill_gradient(
    low = "white",
    high = "steelblue",
    labels = scales::percent_format()
  ) +
  labs(
    title = "Conversion Rate Heatmap",
    x = "Account Tier",
    y = "Industry",
    fill = "Conversion"
  ) +
  theme_minimal()
```
# Industry Conversion
- Above baseline -- blood & biologic and medical device perform stronger than baseline

```{r}
df_clean %>%
  mutate(is_sql = next_stage_c == "SQL") %>%
  group_by(acct_target_industry, is_sql) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(acct_target_industry) %>%
  mutate(pct = count / sum(count))
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```