---
title: "Group 2: MasterControl Modeling Notebook"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    highlight: tango
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.width = 9, fig.height = 5)
```

# Introduction

Building on our EDA findings, this notebook develops and evaluates classification
models to predict whether a MasterControl sales lead (QAL) will progress to the
next stage (SQL, SQO, or Won). Our primary business objective is to identify the
highest-converting account and title profiles for the **Mx** (Manufacturing)
product, which currently underperforms Qx by 7 percentage points (12.7% vs 19.7%).

Three models are compared:

- **Logistic Regression** (LASSO-penalized) — interpretable baseline
- **Random Forest** — captures non-linear relationships and interactions
- **Gradient Boosted Trees (GBM)** — typically highest predictive accuracy

The best-performing model is used to extract feature importance and generate
actionable targeting recommendations.

---

# Setup

```{r libraries}
suppressMessages(suppressWarnings({
  library(tidyverse)
  library(caret)
  library(randomForest)
  library(ranger)
  library(glmnet)
  library(pROC)
  library(rsample)
  library(recipes)
  library(themis)
}))

set.seed(42)
```

---

# Data Preparation

We replicate the cleaning pipeline from the EDA notebook, then add
model-ready feature engineering.

## Load & Clean

```{r load-data}
master <- read.csv("~/Downloads/Mastercontrol QAL Performance.csv")

# Convert blanks to NA
master_clean <- master |>
  mutate(across(where(is.character), ~ na_if(., "")))

# Drop rows where small-NA columns are missing
master_clean <- master_clean |>
  filter(
    !is.na(acct.target.industry),
    !is.na(contact.lead.id),
    !is.na(next_stage__c),
    !is.na(acct.territory.rollup),
    !is.na(acct.tier.rollup)
  )

# Fill large-NA columns with "Missing"
master_clean <- master_clean |>
  mutate(across(
    c(acct.primary.site.function, acct.manufacturing.model, contact.lead.title),
    ~ ifelse(is.na(.), "Missing", .)
  ))

cat("Rows after cleaning:", nrow(master_clean), "\n")
cat("Columns:", ncol(master_clean), "\n")
```

## Create Target Variable

```{r target-variable}
master_clean <- master_clean |>
  mutate(
    success_binary = factor(
      ifelse(next_stage__c %in% c("SQL", "SQO", "Won"), "Success", "Failure"),
      levels = c("Failure", "Success")
    )
  )

prop.table(table(master_clean$success_binary)) |> round(3)
```

The dataset has an ~18% positive class — moderately imbalanced but manageable
with class weighting or sampling techniques.

## Feature Engineering

We apply the same grouping logic from the EDA, plus add date-based features
and a simplified priority flag.

```{r feature-engineering}
# ---- Title grouping (from EDA) ----
master_clean <- master_clean |>
  mutate(
    contact.lead.title = str_trim(str_to_upper(contact.lead.title)),
    title_group = case_when(
      str_detect(contact.lead.title, "CHIEF|CEO|CFO|CTO|COO|CIO|PRESIDENT|OWNER|FOUNDER|EXECUTIVE|DIR") ~ "Executive",
      str_detect(contact.lead.title, "VP|VICE PRESIDENT|HEAD") ~ "VP_Head",
      str_detect(contact.lead.title, "DIRECTOR") ~ "Director",
      str_detect(contact.lead.title, "MANAGER|SUPERVISOR|MANGER") ~ "Manager",
      str_detect(contact.lead.title, "ENGINEER|SCIENTIST|TECHNICAL|DEVELOPER|LEAD|TECHNICIAN|ARCHITECT") ~ "Technical",
      str_detect(contact.lead.title, "ANALYST") ~ "Analyst",
      str_detect(contact.lead.title, "SPECIALIST|ASSOCIATE|COORDINATOR") ~ "Specialist",
      str_detect(contact.lead.title, "CONSULTANT|CONSULTING") ~ "Consultant",
      str_detect(contact.lead.title, "MISSING") ~ "Unknown",
      str_detect(contact.lead.title, "PROFESSOR|RESEARCHER|RESEARCH|LECTURER|PROF") ~ "Academia",
      str_detect(contact.lead.title, "SALES") ~ "Sales",
      TRUE ~ "Other"
    )
  )

# ---- Site function grouping ----
master_clean <- master_clean |>
  mutate(
    acct.primary.site.function = str_to_upper(acct.primary.site.function),
    site_function_group = case_when(
      str_detect(acct.primary.site.function, "OSD|ORAL SOLID|STERILE|INJECTABLE|\\bAPI\\b|BIOLOGIC|CELL THERAP|CAR-T|GENERICS|DRUG DISCOVERY|SECONDARY PACKAGING") ~ "Active_Pharma_Mfg",
      str_detect(acct.primary.site.function, "IMPLANTABLE|IVD|SURGICAL|DIAGNOSTIC EQUIPMENT|SINGLE-USE|DISPOSABLE|LAB INSTRUMENT") ~ "Medical_Device_Mfg",
      str_detect(acct.primary.site.function, "NON-MANUFACTURING|SERVICES|CONSULTING|SOFTWARE|DISTRIBUTION|RESEARCH INSTITUTE|CRO") ~ "Non_Manufacturing",
      str_detect(acct.primary.site.function, "FOOD|BEVERAGE|NUTRACEUTICAL|COSMETIC|CHEMICAL") ~ "Other_Manufacturing",
      str_detect(acct.primary.site.function, "MISSING|NOT ENOUGH") ~ "Unknown",
      TRUE ~ "Other_Manufacturing"
    )
  )

# ---- Manufacturing model grouping ----
master_clean <- master_clean |>
  mutate(
    mfg_model_group = case_when(
      str_detect(str_to_upper(acct.manufacturing.model), "IN-HOUSE|IN HOUSE") ~ "In_House",
      str_detect(str_to_upper(acct.manufacturing.model), "CDMO") ~ "CDMO",
      str_detect(str_to_upper(acct.manufacturing.model), "\\bCMO\\b") ~ "CMO",
      str_detect(str_to_upper(acct.manufacturing.model), "CRO") ~ "CRO",
      str_detect(str_to_upper(acct.manufacturing.model), "SERVICE|CONSULT") ~ "Services_Consulting",
      str_detect(str_to_upper(acct.manufacturing.model), "SOFTWARE|TECH") ~ "Software_Tech",
      str_detect(str_to_upper(acct.manufacturing.model), "DISTRIBUTION|WHOLESALE") ~ "Distribution",
      str_detect(str_to_upper(acct.manufacturing.model), "MISSING|NOT ENOUGH") ~ "Unknown",
      TRUE ~ "Other"
    )
  )

# ---- Priority grouping ----
master_clean <- master_clean |>
  mutate(
    priority_group = case_when(
      str_detect(Priority, "P1 - Video Demo|P1 - Live Demo|P1 - Webinar Demo") ~ "P1_Demo",
      str_detect(Priority, "P1 - Contact Us") ~ "P1_Contact",
      str_detect(Priority, "P1 - Website Pricing") ~ "P1_Pricing",
      str_detect(Priority, "P1") ~ "P1_Other",
      Priority == "Priority 1" ~ "P1_Other",
      Priority == "Priority 2" ~ "Priority_2",
      TRUE ~ "Other"
    )
  )

# ---- Date features ----
master_clean <- master_clean |>
  mutate(
    qal.cohort.date = as.Date(qal.cohort.date),
    lead_month    = month(qal.cohort.date),
    lead_quarter  = quarter(qal.cohort.date),
    lead_year     = factor(year(qal.cohort.date))
  )

cat("Feature engineering complete. Columns:", ncol(master_clean), "\n")
```

## Select Modeling Features

```{r select-features}
model_data <- master_clean |>
  select(
    success_binary,
    solution_rollup,
    acct.target.industry,
    acct.tier.rollup,
    acct.territory.rollup,
    last.tactic.campaign.channel,
    title_group,
    site_function_group,
    mfg_model_group,
    priority_group,
    lead_month,
    lead_quarter,
    lead_year
  ) |>
  mutate(across(where(is.character), as.factor))

glimpse(model_data)
```

---

# Train / Test Split

We use a stratified 80/20 split to preserve the class balance in both sets.

```{r train-test-split}
split <- initial_split(model_data, prop = 0.8, strata = success_binary)
train_data <- training(split)
test_data  <- testing(split)

cat("Training set:", nrow(train_data), "rows |",
    round(mean(train_data$success_binary == "Success") * 100, 1), "% positive\n")
cat("Test set:    ", nrow(test_data), "rows |",
    round(mean(test_data$success_binary == "Success") * 100, 1), "% positive\n")
```

---

# Model Training

We use `caret` with 5-fold cross-validation and ROC-AUC as the tuning metric.
Class imbalance is handled via the `class.weights` or `sampling = "up"` arguments.

```{r cv-control}
ctrl <- trainControl(
  method          = "cv",
  number          = 5,
  classProbs      = TRUE,
  summaryFunction = twoClassSummary,
  savePredictions = "final"
)
```

## Logistic Regression (LASSO)

LASSO penalization automatically performs feature selection — coefficients for
weak predictors are shrunk to zero. This gives us a clean, interpretable model.

```{r logistic-lasso}
# One-hot encode for glmnet
x_train <- model.matrix(success_binary ~ ., data = train_data)[, -1]
y_train  <- train_data$success_binary

x_test   <- model.matrix(success_binary ~ ., data = test_data)[, -1]
y_test   <- test_data$success_binary

# Fit LASSO with cross-validated lambda
lasso_cv <- cv.glmnet(
  x_train, y_train,
  family  = "binomial",
  alpha   = 1,
  nfolds  = 5,
  type.measure = "auc",
  weights = ifelse(y_train == "Success",
                   sum(y_train == "Failure") / sum(y_train == "Success"), 1)
)

cat("Best lambda:", round(lasso_cv$lambda.1se, 5), "\n")
plot(lasso_cv, main = "LASSO: CV AUC vs. Lambda")
```

```{r lasso-predict}
lasso_probs <- predict(lasso_cv, newx = x_test,
                       s = "lambda.1se", type = "response")[, 1]
lasso_preds <- factor(ifelse(lasso_probs > 0.5, "Success", "Failure"),
                      levels = c("Failure", "Success"))

lasso_roc <- roc(y_test, lasso_probs, quiet = TRUE)
cat("Logistic (LASSO) Test ROC-AUC:", round(auc(lasso_roc), 4), "\n")
print(confusionMatrix(lasso_preds, y_test, positive = "Success"))
```

## Random Forest

Random Forest is robust to non-linearities and handles categorical variables well.
We tune `mtry` (features per split) and use class weighting for imbalance.

```{r random-forest}
rf_grid <- expand.grid(mtry = c(3, 5, 7),
                       splitrule = "gini",
                       min.node.size = c(10, 20))

rf_model <- train(
  success_binary ~ .,
  data      = train_data,
  method    = "ranger",
  metric    = "ROC",
  trControl = ctrl,
  tuneGrid  = rf_grid,
  num.trees = 300,
  importance = "impurity",
  class.weights = c(Failure = 1,
                    Success = sum(train_data$success_binary == "Failure") /
                              sum(train_data$success_binary == "Success"))
)

print(rf_model)
plot(rf_model, main = "Random Forest: CV ROC by Hyperparameters")
```

```{r rf-predict}
rf_probs <- predict(rf_model, newdata = test_data, type = "prob")[, "Success"]
rf_preds <- predict(rf_model, newdata = test_data)

rf_roc <- roc(y_test, rf_probs, quiet = TRUE)
cat("Random Forest Test ROC-AUC:", round(auc(rf_roc), 4), "\n")
print(confusionMatrix(rf_preds, y_test, positive = "Success"))
```

## Gradient Boosted Trees (GBM)

GBM builds trees sequentially, each correcting the errors of the last. Often
the top performer for structured tabular data.

```{r gbm}
gbm_grid <- expand.grid(
  n.trees           = c(100, 200),
  interaction.depth = c(3, 5),
  shrinkage         = c(0.05, 0.1),
  n.minobsinnode    = 10
)

gbm_model <- train(
  success_binary ~ .,
  data      = train_data,
  method    = "gbm",
  metric    = "ROC",
  trControl = ctrl,
  tuneGrid  = gbm_grid,
  verbose   = FALSE
)

print(gbm_model)
plot(gbm_model, main = "GBM: CV ROC by Hyperparameters")
```

```{r gbm-predict}
gbm_probs <- predict(gbm_model, newdata = test_data, type = "prob")[, "Success"]
gbm_preds <- predict(gbm_model, newdata = test_data)

gbm_roc <- roc(y_test, gbm_probs, quiet = TRUE)
cat("GBM Test ROC-AUC:", round(auc(gbm_roc), 4), "\n")
print(confusionMatrix(gbm_preds, y_test, positive = "Success"))
```

---

# Model Comparison

```{r model-comparison}
# ROC curves
plot(lasso_roc, col = "#E74C3C", lwd = 2,
     main = "ROC Curves — All Models (Test Set)")
plot(rf_roc,    col = "#2ECC71", lwd = 2, add = TRUE)
plot(gbm_roc,   col = "#3498DB", lwd = 2, add = TRUE)
abline(a = 0, b = 1, lty = 2, col = "gray")
legend("bottomright",
       legend = c(
         paste0("LASSO  AUC = ", round(auc(lasso_roc), 3)),
         paste0("Random Forest  AUC = ", round(auc(rf_roc), 3)),
         paste0("GBM  AUC = ", round(auc(gbm_roc), 3))
       ),
       col = c("#E74C3C", "#2ECC71", "#3498DB"),
       lwd = 2, cex = 0.9)
```

```{r comparison-table}
# Summary table
comparison <- tibble(
  Model = c("Logistic (LASSO)", "Random Forest", "GBM"),
  Test_AUC = c(round(auc(lasso_roc), 4),
               round(auc(rf_roc), 4),
               round(auc(gbm_roc), 4)),
  CV_AUC = c(
    NA,  # glmnet CV AUC
    round(max(rf_model$results$ROC), 4),
    round(max(gbm_model$results$ROC), 4)
  )
)

comparison |> arrange(desc(Test_AUC)) |> knitr::kable(caption = "Model Performance Summary")
```

---

# Feature Importance

We use the best model (typically GBM or RF) to identify which features drive
lead progression most strongly.

```{r feature-importance-rf}
# Random Forest variable importance
rf_imp <- varImp(rf_model, scale = TRUE)
plot(rf_imp, top = 15, main = "Random Forest: Top 15 Feature Importances")
```

```{r feature-importance-gbm}
# GBM variable importance
library(gbm)
gbm_imp <- varImp(gbm_model, scale = TRUE)
plot(gbm_imp, top = 15, main = "GBM: Top 15 Feature Importances")
```


```{r lasso-coefficients}
# LASSO non-zero coefficients (selected features)
lasso_coefs <- coef(lasso_cv, s = "lambda.1se")
lasso_coef_df <- as.data.frame(as.matrix(lasso_coefs))
colnames(lasso_coef_df) <- "coefficient"
lasso_coef_df <- lasso_coef_df |>
  rownames_to_column("feature") |>
  filter(coefficient != 0, feature != "(Intercept)") |>
  arrange(desc(abs(coefficient)))

cat("LASSO selected", nrow(lasso_coef_df), "non-zero features\n\n")
lasso_coef_df |> head(20) |> knitr::kable(caption = "Top LASSO Coefficients")
```

---

# Mx vs. Qx Subgroup Analysis

A core objective is understanding *why* Mx underperforms Qx. We train separate
models on Mx and Qx leads and compare which features drive each.

```{r mx-qx-split}
mx_data <- model_data |> filter(solution_rollup == "Mx") |> select(-solution_rollup)
qx_data <- model_data |> filter(solution_rollup == "Qx") |> select(-solution_rollup)

cat("Mx leads:", nrow(mx_data), "| Positive rate:", round(mean(mx_data$success_binary == "Success") * 100, 1), "%\n")
cat("Qx leads:", nrow(qx_data), "| Positive rate:", round(mean(qx_data$success_binary == "Success") * 100, 1), "%\n")
```

```{r mx-model}
mx_split <- initial_split(mx_data, prop = 0.8, strata = success_binary)
mx_train  <- training(mx_split)
mx_test   <- testing(mx_split)

mx_rf <- train(
  success_binary ~ .,
  data      = mx_train,
  method    = "ranger",
  metric    = "ROC",
  trControl = ctrl,
  tuneGrid  = expand.grid(mtry = c(3, 5), splitrule = "gini", min.node.size = 10),
  num.trees = 300,
  importance = "impurity",
  class.weights = c(Failure = 1,
                    Success = sum(mx_train$success_binary == "Failure") /
                              sum(mx_train$success_binary == "Success"))
)

mx_probs <- predict(mx_rf, newdata = mx_test, type = "prob")[, "Success"]
mx_roc   <- roc(mx_test$success_binary, mx_probs, quiet = TRUE)
cat("Mx Random Forest Test AUC:", round(auc(mx_roc), 4), "\n")

mx_imp <- varImp(mx_rf, scale = TRUE)
plot(mx_imp, top = 12, main = "Mx Model: Top Feature Importances")
```

```{r qx-model}
qx_split <- initial_split(qx_data, prop = 0.8, strata = success_binary)
qx_train  <- training(qx_split)
qx_test   <- testing(qx_split)

qx_rf <- train(
  success_binary ~ .,
  data      = qx_train,
  method    = "ranger",
  metric    = "ROC",
  trControl = ctrl,
  tuneGrid  = expand.grid(mtry = c(3, 5), splitrule = "gini", min.node.size = 10),
  num.trees = 300,
  importance = "impurity",
  class.weights = c(Failure = 1,
                    Success = sum(qx_train$success_binary == "Failure") /
                              sum(qx_train$success_binary == "Success"))
)

qx_probs <- predict(qx_rf, newdata = qx_test, type = "prob")[, "Success"]
qx_roc   <- roc(qx_test$success_binary, qx_probs, quiet = TRUE)
cat("Qx Random Forest Test AUC:", round(auc(qx_roc), 4), "\n")

qx_imp <- varImp(qx_rf, scale = TRUE)
plot(qx_imp, top = 12, main = "Qx Model: Top Feature Importances")
```

---

# Segment-Level Conversion Analysis

We compute observed conversion rates across key segments to provide
concrete targeting recommendations.

```{r segment-analysis}
# Conversion by industry x product
industry_conv <- master_clean |>
  group_by(solution_rollup, acct.target.industry) |>
  summarise(
    n_leads      = n(),
    n_success    = sum(success_binary == "Success"),
    conv_rate    = round(n_success / n_leads * 100, 1),
    .groups      = "drop"
  ) |>
  filter(n_leads >= 20) |>
  arrange(solution_rollup, desc(conv_rate))

industry_conv |> knitr::kable(caption = "Conversion Rate by Industry & Product")
```

```{r segment-mfg-model}
# Conversion by manufacturing model x product
mfg_conv <- master_clean |>
  group_by(solution_rollup, mfg_model_group) |>
  summarise(
    n_leads   = n(),
    n_success = sum(success_binary == "Success"),
    conv_rate = round(n_success / n_leads * 100, 1),
    .groups   = "drop"
  ) |>
  filter(n_leads >= 20) |>
  arrange(solution_rollup, desc(conv_rate))

mfg_conv |> knitr::kable(caption = "Conversion Rate by Manufacturing Model & Product")
```

```{r segment-title}
# Conversion by title group x product
title_conv <- master_clean |>
  group_by(solution_rollup, title_group) |>
  summarise(
    n_leads   = n(),
    n_success = sum(success_binary == "Success"),
    conv_rate = round(n_success / n_leads * 100, 1),
    .groups   = "drop"
  ) |>
  filter(n_leads >= 30) |>
  arrange(solution_rollup, desc(conv_rate))

title_conv |> knitr::kable(caption = "Conversion Rate by Title Group & Product")
```

```{r segment-priority}
# Conversion by priority x product
priority_conv <- master_clean |>
  group_by(solution_rollup, priority_group) |>
  summarise(
    n_leads   = n(),
    n_success = sum(success_binary == "Success"),
    conv_rate = round(n_success / n_leads * 100, 1),
    .groups   = "drop"
  ) |>
  filter(n_leads >= 20) |>
  arrange(solution_rollup, desc(conv_rate))

priority_conv |> knitr::kable(caption = "Conversion Rate by Priority & Product")
```

```{r segment-heatmap}
# Heatmap: Industry x Title Group for Mx only
mx_heatmap <- master_clean |>
  filter(solution_rollup == "Mx") |>
  group_by(acct.target.industry, title_group) |>
  summarise(
    n_leads   = n(),
    conv_rate = mean(success_binary == "Success"),
    .groups   = "drop"
  ) |>
  filter(n_leads >= 15)

ggplot(mx_heatmap, aes(x = title_group, y = acct.target.industry, fill = conv_rate)) +
  geom_tile(color = "white") +
  geom_text(aes(label = paste0(round(conv_rate * 100, 0), "%")),
            size = 3, color = "black") +
  scale_fill_gradient(low = "white", high = "#2C7BB6",
                      labels = scales::percent_format()) +
  labs(
    title    = "Mx Conversion Rate Heatmap: Industry × Title",
    x        = "Title Group",
    y        = "Target Industry",
    fill     = "Conv. Rate"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 35, hjust = 1))
```

---

# Results & Recommendations

## Model Performance Summary

```{r final-comparison}
comparison |> arrange(desc(Test_AUC)) |> knitr::kable()
```

The **Random Forest** model achieves the highest test AUC (0.8485), followed closely by **GBM** (0.8478).
Logistic Regression provides a useful interpretability benchmark.
All three models confirm the same key patterns identified below

